### 2023-04-19 manuscript.Rmd
Update text to match revision

authornote: |  
  Number of pages: ???

  Number of figures: 4

  Number of words: ??? (abstract), ??? (introduction), ??? (discussion)

  We would like to thank Nele Langosch for assistance with the preparation of materials and data collection, Olaf Dimigen and the Abdel Rahman Lab for Neurocognitive Psychology for insightful discussions, Guido Kiecker for technical support, as well as Guillaume Thierry and one anonymous reviewer for their comments on a previous version of the paper.
  This work was supported by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany’s Excellence Strategy (EXC 2002/1 "Science of Intelligence") project number 390523135 as well as grants AB277-5 and AB277-6 to R.A.R.

  The authors declare no competing financial interests.

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

documentclass     : "apa6"
classoption       : "man,donotrepeattitle"
fontsize          : "10pt"

@@

header-includes:
  - \geometry{a4paper}
  - \raggedbottom
  - \usepackage{orcidlink}
  - \usepackage[all]{nowidow}
  - \captionsetup[figure]{belowskip=10pt}
  - \captionsetup[figure]{font={small,stretch=1},textfont=normalfont}
  - \captionsetup[figure]{labelfont=bf,labelsep=period,name=Figure}
  - \captionsetup[table]{labelfont=bf,labelsep=period,textfont=bf}
  - \usepackage[export]{adjustbox}
  - \usepackage{makecell}
  - \usepackage{setspace}
---

@@

# Set options
options(readr.show_col_types = FALSE)
emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
opts_chunk$set(
  fig.pos = "tp",
  fig.width = 12,
  message = FALSE,
  out.width = "100%",
  warning = FALSE
)

@@

Does our perception of an object change once we discover what function it serves?
We showed human participants (*n* = 48) pictures of unfamiliar objects either together with keywords matching their function, leading to semantically informed perception, or together with non-matching keywords, resulting in [uninformed]{color="blue"} perception.
We measured event-related potentials (ERPs) to investigate at which stages in the visual processing hierarchy these two types of object perception differed from one another.
We found that semantically informed as compared to [uninformed]{color="blue"} perception was associated with larger amplitudes in the N170 component (150--200 ms), reduced amplitudes in the N400 component (400--700 ms), and a late decrease in alpha/beta band power.
When the same objects were presented once more without any information, the N400 and event-related power effects persisted, and we also observed enlarged amplitudes in the P1 component (100--150 ms) in response to objects for which semantically informed perception had taken place.
Consistent with previous work, this suggests that obtaining semantic information about previously unfamiliar objects alters aspects of their lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component, event-related power).
Our study is the first to show that such effects [occur instantly after semantic information has been provided for the first time, without requiring  extensive learning.]{color="blue"}

*Key words*: objects; semantic knowledge; visual perception; event-related potentials

@@

Does our perception of an object change once we discover what function it serves?
This question speaks to the long-standing debate about the cognitive (im)penetrability of perception by higher-level capacities such as semantic knowledge or language [@pylyshyn1999].
According to one view, these cognitive capacities kick in only after the retinal input has been processed by a specialized module for visual perception [@firestone2016; @fodor1983].
This module is supposed to be encapsulated from higher-level inputs and processes visual information in a purely feed-forward fashion, progressing from lower areas with small receptive field sizes to areas representing increasingly complex shapes and, eventually, whole objects [@dicarlo2012].

The cognitive impenetrability hypothesis is challenged by the alternative view that perception dynamically interacts with different cognitive aspects from early on [@churchland1994; @ahissar2004; @yuille2006; @clark2013; @lupyan2015; @thierry2016; @teufel2017; @lupyan2020].
On the behavioral level, this is supported by psychological studies that found differences in ratings, detection rates, or reaction times for visual stimuli depending on their emotional [e.g., @phelps2016], motivational [e.g., @balcetis2010], linguistic [e.g., @slivac2021] or semantic [e.g., @gauthier2003] content.
[However, some studies have received legitimate criticism for a number of reasons [@firestone2016].
For example, tentative high-level effects (e.g., the emotional difference between a smiling and a neutral facial expression) were sometimes confounded with low-level visual differences (e.g., the difference in visual contrast introduced by one face showing their teeth and the other one not).
As another example, studies that relied on behavioral measures such as reaction times were not able to discern effects of higher-level cognition on perception from effects that occur in later, post-perceptual stages of processing [e.g., easier recognition from memory for semantically meaningful stimuli, despite no change in how they are perceived visually\; @firestone2016].]{color="blue"}

On the physiological level, event-related potentials (ERPs) measured from the human EEG can mitigate most of these concerns:
Their temporal resolution makes it possible to probe how early one can detect influences of high-level (e.g., semantic) information [@athanasopoulos2020].
To this end, participants in previous studies by our lab and others learned to associate visual objects with different amounts of semantic information.
After learning, an orthogonal task was used to compare ERPs in response to objects learned with versus without semantic information.
This revealed differences not only in late ERP components associated with semantic processing (i.e., the N400 component) but also in the visual P1 component [@abdelrahman2008; @maier2019; @samaha2018; @weller2019].
The early peak of this component and its source in the occipital cortex [@abdelrahman2008] point to an immediate effect of semantic knowledge on visual perception.
It is less clear, however, if this tentative top-down effect of semantic cognition on perception acts in an online fashion (i.e., directly modulating perceptual processing), or more indirectly, by altering stored visual [memory]{color="blue"} representations over the course of learning, which would then get reactivated once the object is reencountered later on.
[The first account (i.e., an online top-down effect) would be in accordance with predictive coding theories that posit that higher-level areas influence ongoing perceptual processing by sending predictions down to lower-level areas [e.g., @rao1999; @friston2009; @clark2013; @lupyan2015].
The second account (i.e., re-activation of stored memory representations) would argue against such online effects, positing that differences in ERPs emerge merely due to comparisons with visual memory representations or prototypes that have been built up over the course of learning [e.g., @palmeri2008].]{color="blue"}

To answer this question, we measured ERPs in response to unfamiliar objects directly while participants gained a semantic understanding of their function.
We presented half of the objects together with matching keywords, thus allowing participants to understand what kind of object they were viewing, and the other half together with non-matching keywords, [thus preventing participants from understanding what kind of object they were viewing]{color="blue"}.
We then presented the same objects again to test for downstream effects of semantic information.
We examined the influence of semantic information on ERPs associated with lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component).
We hypothesized that semantic information would have instant effects on visual perception, by which we mean (a) that these effects can already be observed on the same trial as the semantic information is being presented for the first time, and (b) that these effects are found not only in later, higher-level cognition-related ERP components (i.e., N400), but also in earlier, perception-related ERP components (i.e., P1 and/or N170).
[Observing these effects immediately after participants have received the relevant semantic information, instead of after an extensive learning phase as in previous studies, would speak for theories that posit an online influence of higher-level cognitive areas on lower-level visual perception [e.g., predictive coding\; @friston2009].]{color="blue"}
We furthermore conducted an exploratory time-frequency analysis to test for effects on event-related power which, unlike ERP effects, do not need to be tightly phase-locked to stimulus onset.

@@

Participants were 48 German native speakers (31 female, 17 male) with a mean age of 23.5 years (range 18--32 years) and no history of psychological disorder or treatment.
No a priori power analysis was carried out.
All participants were right-handed according to the Edinburgh inventory [@oldfield1971] and reported normal or corrected-to-normal vision.
They provided written informed consent before starting the experiment and received a monetary compensation of €8 per hour for participating.

@@

Stimuli consisted of 240 grayscale photographs of real-world objects.
Of these, 120 stimuli were well-known everyday objects (e.g., a bicycle, a toothbrush).
These served as filler stimuli of no interest.
The other 120 stimuli were rare objects presumed to be unfamiliar to the majority of participants (e.g., a galvanometer, an udu drum; [see Supplementary Table 1]{color="blue"}).
All stimuli were presented on a light blue background with a size of 207 × 207 pixels on a 19-inch LCD monitor with a resolution of 1,280 × 1,024 pixels and a refresh rate of 75 Hz.
At a standardized viewing distance of 90 cm, the images subtended approximately 3.9 degrees of participants' horizontal and vertical visual angle.

For each unfamiliar object, we created a pair of German keywords (a noun and a verb), describing the typical function or use of the object in a way that could be related to its visual features and their configuration (e.g., *Stromstärke, messen* [electric current, measuring]; *Tonpott, trommeln* [clay pot, drumming]; [see Supplementary Table 1]{color="blue"}).
As our central experimental manipulation, half of the objects were presented together with keywords that matched their respective function, whereas the other half of the objects were presented together with non-matching keywords [that would have matched another object]{color="blue"}.
The matching keywords were expected to induce semantically informed perception, that is, participants suddenly understanding what kind of object they were viewing.
The non-matching keywords were expected to prevent such an understanding and keep the perception of the object semantically [uninformed]{color="blue"}.
All participants saw each unfamiliar object with only one type of keywords (matching or non-matching).
This assignment of keywords to objects was counterbalanced across participants so that each object was presented with matching keywords (leading to semantically informed perception) and non-matching keywords (leading to [uninformed]{color="blue"} perception) to an equal number of participants.
The experiment was programmed and displayed using Presentation® software (Neurobehavioral Systems, Inc., Berkeley, CA, www.neurobs.com).

@@

[As a manipulation check, we ran an online rating study where we presented 10 German speakers (3 female, 7 male, mean age = 25.3 years, range 20--35 years; none took part in the main study) with all 240 visual objects in random order and asked them to generate their own keywords that would describe the presumed function of the object.
We used latent semantic analysis [@gunther2019] with a word2vec embedding (deepset GmbH, Berlin, Germany, www.deepset.ai/german-word-embeddings) pre-trained on the German Wikipedia to estimate semantic distances between these participant-generated keywords and the keywords used in our main EEG experiment (Supplementary Figures 1 and 2).
In brief, it was substantially easier for participants to come with correct descriptions of familiar objects (mean cosine distance ± standard error = `r filter(rating_stats$means, fam_condition == "familiar")$emmean` ± `r filter(rating_stats$means, fam_condition == "familiar")$SE`) than for unfamiliar objects (`r filter(rating_stats$means, fam_condition == "matching")$emmean` ± `r filter(rating_stats$means, fam_condition == "matching")$SE`), $t$(`r print_dof(filter(rating_stats$contrasts, contrast == "familiar - matching")$df)`) = `r filter(rating_stats$contrasts, contrast == "familiar - matching")$t_value`, $p_\text{corr}$ `r print_p(filter(rating_stats$contrasts, contrast == "familiar - matching")$p_value)` (linear mixed-effects model).
In fact, this similarity between participant-generated keywords for the unfamiliar objects and the object-matching keywords presented in the main experiment was only slightly higher than the similarity between participant-generated keywords and the object-non-matching keywords presented in the main experiment (`r filter(rating_stats$means, fam_condition == "non-matching")$emmean` ± `r filter(rating_stats$means, fam_condition == "non-matching")$SE`), $t$(`r print_dof(filter(rating_stats$contrasts, contrast == "matching - non-matching")$df)`) = `r filter(rating_stats$contrasts, contrast == "matching - non-matching")$t_value`, $p_\text{corr}$ `r print_p(filter(rating_stats$contrasts, contrast == "matching - non-matching")$p_value)` (linear mixed-effects model), indicating that it was difficult for participants to know or guess the correct function of the unfamiliar objects.]{color="blue"}

@@

The main EEG experiment consisted of three phases (see Figure 1A).
In the *pre-insight* phase, after written informed consent had been obtained and the EEG had been prepared, all 240 familiar and unfamiliar objects were presented once in random order and without any keywords.
Each trial consisted of a fixation cross presented in the middle of the screen for 0.5 s, followed by the presentation of the object until participants made a response or until a time out after 3 s.
The inter-trial interval was 0.5 s and participants took a self-timed break after each block of 60 objects.
The task, which was kept the same across all three phases of the experiment, was to classify each object using one of four response alternatives:
(a) *Ich weiß, was das ist, oder habe eine starke Vermutung* [I know what this is or have a strong assumption],
(b) *Ich habe eher eine Vermutung, was das ist* [I rather have an assumption what this is],
(c) *Ich habe eher keine Vermutung, was das ist* [I rather have no assumption what this is], or
(d) *Ich weiß nicht, was das ist, und habe auch keine Vermutung* [I don’t know what this is and have no assumption].
Participants were asked to respond as quickly and as accurately as possible by pressing one out of four buttons with the index or middle finger of their left or right hand, respectively.
The mapping of the rating scale to the four buttons (left to right or right to left) was counterbalanced across participants.

@@

```{r, fig1, cache=!run$figures, fig.height=12, fig.cap="(ref:fig1-caption)"}
# Create Figure 1
plot_fig1(misc_dir, evokeds, config, channel_locations, models)
```

(ref:fig1-caption) Experimental design and ERP results [contrasting semantically informed perception and uninformed perception]{color="blue"}.
***A***, In the pre-insight phase, participants were presented with 120 unfamiliar objects and indicated if they knew what kind of object they were viewing.
In the insight phase, half of these objects were presented with matching keywords (in red color for illustration), leading to semantically informed perception, and the other half with non-matching keywords (in blue color for illustration), leading to [uninformed]{color="blue"} perception.
In the post-insight phase, the same objects were presented again without keywords.
***B***, ***C***, ***D***, ERP waveforms and scalp topographies for the P1 component (***B***), for the N170 component (***C***), and for the N400 component (***D***) for objects with semantically informed perception versus [uninformed]{color="blue"} perception within the three different phases.
Semantically informed perception was associated with more negative amplitudes in the N170 component during the insight phase, less negative amplitudes in the N400 component during the insight and post-insight phases, and more positive amplitudes in the P1 component during the post-insight phase.
[Waveform plots show the ERP amplitudes averaged across channels in the regions of interest (P1: PO3, PO4, POz, O1, O2, Oz; N170: P7, P8, PO7, PO8, PO9, PO10; N400: C1, C2, Cz, CP1, CP2, CPz; see black dots in the scalp topographies).]{color="blue"}
Colored ribbons around the ERP waveforms show ± 1 standard error of the mean across participants.
[Topographies show the difference in ERP amplitudes at all channels on the scalp, averaged across the time windows of interest (P1: 100--150 ms, N170: 150--200 ms, N400: 400--700 ms; see gray areas in the ERP waveforms).]{color="blue"}
Ampl. = amplitude.
\* $p_\text{corr} < .05$. \*\*\* $p_\text{corr} < .001$.

In the *insight* phase, the 120 unfamiliar objects were presented for a second time, now preceded either by matching keywords (leading to semantically informed perception) or by non-matching keywords (leading to [uninformed]{color="blue"} perception).
Each trial consisted of a fixation cross presented for 0.5 s, followed by the presentation of the keywords for 2.5 s.
Then, an asterisk was presented in the middle of the screen for another 0.5 s, followed by the presentation of the object until a response was made or until a time out after 3 s.
The objects were presented in blocks of 30 trials so that within each block there were 15 objects from each of the two experimental conditions and so that objects were heterogeneous in terms of their shape, visual complexity, and functional category (e.g., medical devices, musical instruments).

In the *post-insight* phase, the unfamiliar objects were presented for a third time, with the same trial structure as in the pre-insight phase, that is, without any keywords.
The insight and post-insight phases were presented in an interleaved fashion so that after the presentation of one block of 30 objects in the insight phase (with keywords), participants took a self-timed break and continued with the same block of 30 objects in the post-insight phase (without keywords) before moving on to the next block consisting of 30 different objects.
They continued like this until all four blocks were completed in both phases.
In total, the experiment consisted of 480 trials (120 familiar objects in the pre-insight phase and 120 unfamiliar objects in the pre-insight, insight, and post-insight phases).
Participants took approximately 35 minutes to complete the entire experiment.

@@

[We used participants' behavioral responses to verify our experimental manipulation and to assign each object to a semantic condition.
First, we used participants' responses from the pre-insight phase (when objects were presented for the first time, without keywords) to make sure that objects were indeed unfamiliar to them.
That is, we excluded any objects for which participants responded with "I know what this is" in the pre-insight phase.
Next, we used participants' responses from the insight phase (when objects were presented for the second time, preceded by keywords) to assign the remaining objects to different semantic conditions: Semantically informed perception, uninformed perception, and unsuccessfully informed perception (see Figure 1A and 2A).
The semantically informed condition consisted of objects that were presented with matching keywords and for which the participant responded with knowing what the object was or having an assumption.
The uninformed condition consisted of objects that were presented with non-matching keywords and for which the participant responded with not knowing what the object was or having rather no assumption.
The unsuccessfully informed condition consisted of objects that were presented with matching keywords but for which the participant responded with not knowing what the object was or having rather no assumption.
The few objects that were presented with non-matching keywords but for which participants responded with knowing what the object was or having an assumption (i.e., "unsuccessfully uninformed" perception) were dropped from all further analyses.
This assignment of objects to semantic conditions (informed, uninformed, or unsuccessfully informed) was carried over from the insight phase to the other two phases (pre-insight and post-insight).
This allowed us to test, on the one hand, if the objects differed in important aspects even before any keywords were presented (pre-insight phase) and, on the other hand, if the semantic information acquired in the insight phase had any down-stream effects on the subsequent perception of the objects (post-insight phase).]{color="blue"}

@@

For statistical analysis of event-related power, we conducted cluster-based permutation tests [@maris2007], separately for each of the three phases of the experiment (pre-insight, insight, and post-insight).
First, we averaged trials belonging to the same condition and then subtracted these average responses from one another to compute the difference between the semantically informed condition and [each of two control conditions (uninformed or unsuccessfully informed)]{color="blue"} for each participant.
We then conducted one-sample $t$ tests in a mass-univariate fashion and grouped significant results (cluster-forming threshold $p < .05$) into clusters if they occurred at neighboring time points, frequencies, or channels.
Neighboring channels were defined using the Delaunay triangulation based on 2D electrode locations as implemented in MNE-Python.
To obtain family-wise error-corrected $p$ values at the cluster level, we compared the cluster mass (i.e., the sum of the $t$ values) of each observed cluster to an empirical null distribution of cluster masses obtained from 5,000 permutations with random sign flips.
Clusters were considered to be statistically significant if their mass exceeded the `r (1 - p_cluster) * 100`^th^ percentile of this distribution (i.e., cluster-level threshold $p < .05$, Bonferroni-corrected for three phases of the experiment).

@@

The EEG data are available upon reasonable request from the corresponding author because we had not asked participants for their consent to make the data publicly available.
[The experimental stimuli (object images and keywords) and]{color="blue"} the code for data analysis are available at [https://doi.org/10.17605/osf.io/uksbc](https://doi.org/10.17605/osf.io/uksbc).

>>>>

authornote: |  
  Alexander Enge \orcidlink{0000-0003-0100-2297} [0000-0003-0100-2297](https://orcid.org/0000-0003-0100-2297), Rasha Abdel Rahman \orcidlink{0000-0002-8438-1570} [0000-0002-8438-1570](https://orcid.org/0000-0002-8438-1570).

  We would like to thank Nele Langosch for assistance with the preparation of materials and data collection, Olaf Dimigen and the Abdel Rahman Lab for Neurocognitive Psychology for insightful discussions, Guido Kiecker for technical support, as well as Guillaume Thierry and one anonymous reviewer for their comments on a previous version of the paper.
  This work was supported by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy (EXC 2002/1 "Science of Intelligence") project number 390523135 as well as grants AB277-5 and AB277-6 to R.A.R.
  The authors declare no competing financial interests.

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

@@

documentclass     : "apa6"
classoption       : "doc,donotrepeattitle"
fontsize          : "10pt"

@@

header-includes:
  - \geometry{a4paper}
  - \raggedbottom
  - \usepackage{orcidlink}
  - \usepackage{tcolorbox}
  - \usepackage[all]{nowidow}
  - \captionsetup[figure]{aboveskip=10pt}
  - \captionsetup[figure]{font={scriptsize,stretch=1},textfont=normalfont}
  - \captionsetup[figure]{labelfont=bf,labelsep=period,name=Figure}
  - \captionsetup[table]{labelfont=bf,labelsep=period,textfont=bf}
  - \usepackage[export]{adjustbox}
  - \usepackage{makecell}
  - \usepackage{setspace}
---

@@

# Set options
options(readr.show_col_types = FALSE)
emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
opts_chunk$set(
  fig.align = "center",
  fig.pos = "tp",
  fig.width = 12,
  message = FALSE,
  out.width = "100%",
  warning = FALSE
)

@@

\smallskip\begin{center}\parbox{12.7cm}{
Does our perception of an object change once we discover what function it serves?
We showed human participants (\textit{n} = 48, 31 females and 17 males) pictures of unfamiliar objects either together with keywords matching their function, leading to semantically informed perception, or together with non-matching keywords, resulting in uninformed perception.
We measured event-related potentials (ERPs) to investigate at which stages in the visual processing hierarchy these two types of object perception differed from one another.
We found that semantically informed as compared to uninformed perception was associated with larger amplitudes in the N170 component (150--200 ms), reduced amplitudes in the N400 component (400--700 ms), and a late decrease in alpha/beta band power.
When the same objects were presented once more without any information, the N400 and event-related power effects persisted, and we also observed enlarged amplitudes in the P1 component (100--150 ms) in response to objects for which semantically informed perception had taken place.
Consistent with previous work, this suggests that obtaining semantic information about previously unfamiliar objects alters aspects of their lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component, event-related power). 
Our study is the first to show that such effects occur instantly after semantic information has been provided for the first time, without requiring extensive learning.
}\end{center}\bigskip

@@

Does our perception of an object change once we discover what function it serves?
This question speaks to the long-standing debate about the cognitive (im)penetrability of perception by higher-level capacities such as semantic knowledge or language.
According to one view, these cognitive capacities kick in only after the retinal input has been processed by a specialized module for visual perception [@fodor1983].
This module is supposed to be encapsulated from higher-level inputs and processes visual information in a feed-forward fashion, progressing from lower to higher areas representing increasingly complex shapes and, eventually, whole objects [@dicarlo2012].
This cognitive impenetrability hypothesis is challenged by another view, namely predictive coding theories that posit that higher-level areas influence ongoing perceptual processing early on by sending predictions down to lower-level areas [@churchland1994; @ahissar2004; @friston2009; @yuille2006; @clark2013; @lupyan2015; @thierry2016; @teufel2017; @lupyan2020].
This view is supported by differences in ratings, detection rates, or reaction times for visual stimuli depending on their emotional [e.g., @phelps2016], linguistic [e.g., @slivac2021] or semantic [e.g., @gauthier2003] content.
However, some of these studies received legitimate criticism, e.g., because comparisons between critical conditions included the confound of additionally comparing different visual stimuli, or for not being able to distinguish between perceptual or post-perceptual loci of tentative top-down effects based on behavioral measures [@firestone2016].

Here we employ event-related potentials (ERPs) measured from the human EEG to mitigate most of these concerns: The temporal resolution of ERPs makes it possible to probe how early influences of high-level (e.g., semantic) information can be detected [@athanasopoulos2020].
When comparing objects after learning different amounts of semantic information about them, previous studies by our lab and others revealed differences in late ERP components associated with semantic processing (N400 component) and also in the visual P1 component [@abdelrahman2008; @samaha2018; @maier2019; @weller2019].
The early peak of the P1 and its source in the occipital cortex [@abdelrahman2008] point to an early effect of semantic knowledge on visual perception.
It is less clear, however, if this effect acts in an online fashion (i.e., directly modulating perceptual processing, in accordance with predictive coding theories), or more indirectly, by altering stored visual object representations over the course of learning, which would then get reactivated once the object is reencountered later on [i.e., reflecting offline differences in visual object representations or prototypes that have been built up over the course of learning\; @palmeri2008].

To answer this question, we measured ERPs in response to unfamiliar objects directly while participants gained a semantic understanding of their function.
We presented half of the objects with matching keywords, allowing participants to understand what kind of object they were viewing, and the other half with non-matching keywords, preventing participants from understanding what kind of object they were viewing.
We then presented the same objects again to test for downstream effects of semantic information.
We examined the influence of semantic information on ERPs associated with lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component).
We hypothesized that semantic information would have instant effects on visual perception, by which we mean (a) that these effects can be observed on the same trial as the semantic information is being presented for the first time, and (b) that these effects are found not only in later, higher-level cognition-related ERP components (N400), but also in earlier, perception-related ERP components (P1 and/or N170), which would speak for predictive coding theories.
We furthermore conducted an exploratory time-frequency analysis to test for effects on event-related power which, unlike ERP effects, do not need to be tightly phase-locked to stimulus onset.

@@

Participants were 48 German native speakers (31 females and 17 males) with a mean age of 23.5 years (range 18--32 years) and no history of psychological disorder or treatment.
No a priori power analysis was carried out.
All participants were right-handed according to the Edinburgh inventory [@oldfield1971] and reported normal or corrected-to-normal vision.
They provided written informed consent before starting the experiment and received a monetary compensation of €8 per hour for participating.

@@

Stimuli consisted of 240 grayscale photographs of real-world objects.
Of these, 120 stimuli were well-known everyday objects (e.g., a bicycle, a toothbrush).
These served as filler stimuli of no interest. The other 120 stimuli were rare objects presumed to be unfamiliar to most participants (e.g., a galvanometer, an udu drum; see Supplementary Table 1 at [https://doi.org/10.17605/osf.io/uksbc](https://doi.org/10.17605/osf.io/uksbc)).
All stimuli were presented on a light blue background with a size of 207 × 207 pixels on a 19-inch LCD monitor with a resolution of 1,280 × 1,024 pixels and a refresh rate of 75 Hz. At a standardized viewing distance of 90 cm, the images subtended approximately 3.9 degrees of participants' horizontal and vertical visual angle.
For each unfamiliar object, we created a pair of German keywords (a noun and a verb), describing the typical function or use of the object in a way that could be related to its visual features and their configuration (e.g., Stromstärke, messen [electric current, measuring]; Tonpott, trommeln [clay pot, drumming]; see Supplementary Table 1).
As our central experimental manipulation, half of the objects were presented together with keywords that matched their respective function, whereas the other half of the objects were presented together with non-matching keywords (which would have matched a different object).
The matching keywords were expected to induce semantically informed perception, that is, participants suddenly understanding what kind of object they were viewing.
The non-matching keywords were expected to prevent such an understanding and keep the perception of the object semantically uninformed.
All participants saw each unfamiliar object with only one type of keywords (matching or non-matching).
This assignment of keywords to objects was counterbalanced across participants so that each object was presented with matching keywords (leading to semantically informed perception) and non-matching keywords (leading to uninformed perception) to an equal number of participants.
The experiment was programmed and displayed using Presentation® software (Neurobehavioral Systems, Inc., Berkeley, CA, www.neurobs.com).

@@

As a manipulation check, we ran an online rating study where we presented 10 German speakers (3 female, 7 male, mean age = 25.3 years, range 20--35 years; none took part in the main study) with all 240 visual objects in random order and asked them to generate their own keywords that would describe the presumed function of the object.
We used latent semantic analysis [@gunther2019] with a word2vec embedding (deepset GmbH, Berlin, Germany, [www.deepset.ai/german-word-embeddings](www.deepset.ai/german-word-embeddings)) pre-trained on the German Wikipedia to estimate semantic distances between these participant-generated keywords and the keywords used in our main EEG experiment (see Supplementary Figures 1 and 2).
In brief, it was substantially easier for participants to come with correct descriptions of familiar objects (mean cosine distance ± standard error = `r filter(rating_stats$means, fam_condition == "familiar")$emmean` ± `r filter(rating_stats$means, fam_condition == "familiar")$SE`) than for unfamiliar objects (`r filter(rating_stats$means, fam_condition == "matching")$emmean` ± `r filter(rating_stats$means, fam_condition == "matching")$SE`), $t$(`r print_dof(filter(rating_stats$contrasts, contrast == "familiar - matching")$df)`) = `r filter(rating_stats$contrasts, contrast == "familiar - matching")$t_value`, $p_\text{corr}$ `r print_p(filter(rating_stats$contrasts, contrast == "familiar - matching")$p_value)` (linear mixed-effects model).
In fact, this similarity between participant-generated keywords for the unfamiliar objects and the object-matching keywords presented in the main experiment was only slightly higher than the similarity between participant-generated keywords and the object-non-matching keywords presented in the main experiment (`r filter(rating_stats$means, fam_condition == "non-matching")$emmean` ± `r filter(rating_stats$means, fam_condition == "non-matching")$SE`), $t$(`r print_dof(filter(rating_stats$contrasts, contrast == "matching - non-matching")$df)`) = `r filter(rating_stats$contrasts, contrast == "matching - non-matching")$t_value`, $p_\text{corr}$ `r print_p(filter(rating_stats$contrasts, contrast == "matching - non-matching")$p_value)` (linear mixed-effects model), indicating that it was difficult for participants to know or guess the correct function of the unfamiliar objects.

@@

The main EEG experiment consisted of three phases (see Figure 1A).
In the *pre-insight* phase, after written informed consent had been obtained and the EEG had been prepared, all 240 familiar and unfamiliar objects were presented once in random order and without any keywords.
Each trial consisted of a fixation cross presented in the middle of the screen for 0.5 s, followed by the presentation of the object until participants made a response or until a time out after 3 s.
The inter-trial interval was 0.5 s and participants took a self-timed break after each block of 60 objects.
The task, which was kept the same across all three phases of the experiment, was to classify each object using one of four response alternatives:
(a) *Ich weiß, was das ist, oder habe eine starke Vermutung* [I know what this is or have a strong assumption],
(b) *Ich habe eher eine Vermutung, was das ist* [I rather have an assumption what this is],
(c) *Ich habe eher keine Vermutung, was das ist* [I rather have no assumption what this is], or
(d) *Ich weiß nicht, was das ist, und habe auch keine Vermutung* [I don't know what this is and have no assumption].
Participants were asked to respond as quickly and as accurately as possible by pressing one out of four buttons with the index or middle finger of their left or right hand, respectively.
The mapping of the rating scale to the four buttons (left to right or right to left) was counterbalanced across participants.

@@

```{r, fig1, cache=!run$figures, fig.height=12, fig.cap="(ref:fig1-caption)"}
# Create Figure 1
plot_fig1(materials_dir, evokeds, config, channel_locations, models)
```

(ref:fig1-caption) Experimental design and ERP results contrasting semantically informed perception and uninformed perception.
***A***, In the pre-insight phase, participants were presented with 120 unfamiliar objects and indicated if they knew what kind of object they were viewing.
In the insight phase, half of these objects were presented with matching keywords (in red color for illustration), leading to semantically informed perception, and the other half with non-matching keywords (in blue color for illustration), leading to uninformed perception.
In the post-insight phase, the same objects were presented again without keywords.
***B***, ***C***, ***D***, ERP waveforms and scalp topographies for the P1 component (***B***), for the N170 component (***C***), and for the N400 component (***D***) for objects with semantically informed perception versus uninformed perception within the three different phases.
Semantically informed perception was associated with more negative amplitudes in the N170 component during the insight phase, less negative amplitudes in the N400 component during the insight and post-insight phases, and more positive amplitudes in the P1 component during the post-insight phase.
Waveform plots show the ERP amplitudes averaged across channels in the regions of interest (P1: PO3, PO4, POz, O1, O2, Oz; N170: P7, P8, PO7, PO8, PO9, PO10; N400: C1, C2, Cz, CP1, CP2, CPz; see black dots in the scalp topographies).
Colored ribbons around the ERP waveforms show ± 1 standard error of the mean across participants.
Topographies show the difference in ERP amplitudes at all channels on the scalp, averaged across the time windows of interest (P1: 100--150 ms, N170: 150--200 ms, N400: 400--700 ms; see gray areas in the ERP waveforms).
Ampl. = amplitude.
\* $p_\text{corr} < .05$. \*\*\* $p_\text{corr} < .001$.

In the *insight* phase, the 120 unfamiliar objects were presented for a second time, now preceded either by matching keywords (leading to semantically informed perception) or by non-matching keywords (leading to uninformed perception).
Each trial consisted of a fixation cross presented for 0.5 s, followed by the presentation of the keywords for 2.5 s.
Then, an asterisk was presented in the middle of the screen for another 0.5 s, followed by the presentation of the object until a response was made or until a time out after 3 s.
The objects were presented in blocks of 30 trials so that within each block there were 15 objects from each of the two experimental conditions and so that objects were heterogeneous in terms of their shape, visual complexity, and functional category (e.g., medical devices, musical instruments).

In the *post-insight* phase, the unfamiliar objects were presented for a third time, with the same trial structure as in the pre-insight phase, that is, without any keywords.
The insight and post-insight phases were presented in an interleaved fashion so that after the presentation of one block of 30 objects in the insight phase (with keywords), participants took a self-timed break and continued with the same block of 30 objects in the post-insight phase (without keywords) before moving on to the next block consisting of 30 different objects.
They continued like this until all four blocks were completed in both phases.
In total, the experiment consisted of 480 trials (120 familiar objects in the pre-insight phase and 120 unfamiliar objects in the pre-insight, insight, and post-insight phases).
Participants took approximately 35 minutes to complete the experiment.

@@

We used participants' behavioral responses to verify our experimental manipulation and to assign each object to a semantic condition.
First, we used participants' responses from the pre-insight phase (when objects were presented for the first time, without keywords) to make sure that objects were indeed unfamiliar to them.
That is, we excluded any objects for which participants responded with "I know what this is" in the pre-insight phase.
Next, we used participants' responses from the insight phase (when objects were presented for the second time, preceded by keywords) to assign the remaining objects to different semantic conditions: Semantically informed perception, uninformed perception, and unsuccessfully informed perception.
The semantically informed condition consisted of objects that were presented with matching keywords and for which the participant responded with knowing what the object was or having an assumption.
The uninformed condition consisted of objects that were presented with non-matching keywords and for which the participant responded with not knowing what the object was or having rather no assumption.
The unsuccessfully informed condition consisted of objects that were presented with matching keywords but for which the participant responded with not knowing what the object was or having rather no assumption.
This assignment of objects to semantic conditions (informed, uninformed, or unsuccessfully informed) was carried over from the insight phase to the other two phases (pre-insight and post-insight).
This allowed us to test, on the one hand, if the objects differed in important aspects even before any keywords were presented (pre-insight phase) and, on the other hand, if the semantic information acquired in the insight phase had any down-stream effects on the subsequent perception of the objects (post-insight phase).

@@

For statistical analysis of event-related power, we conducted cluster-based permutation tests [@maris2007], separately for each of the three phases of the experiment (pre-insight, insight, and post-insight).
First, we averaged trials belonging to the same condition and then subtracted these average responses from one another to compute the difference between the semantically informed condition and each of two control conditions (uninformed or unsuccessfully informed) for each participant.
We then conducted one-sample $t$ tests in a mass-univariate fashion and grouped significant results (cluster-forming threshold $p < .05$) into clusters if they occurred at neighboring time points, frequencies, or channels.
Neighboring channels were defined using the Delaunay triangulation based on 2D electrode locations as implemented in MNE-Python.
To obtain family-wise error-corrected $p$ values at the cluster level, we compared the cluster mass (i.e., the sum of the $t$ values) of each observed cluster to an empirical null distribution of cluster masses obtained from 5,000 permutations with random sign flips.
Clusters were considered to be statistically significant if their mass exceeded the `r (1 - p_cluster) * 100`^th^ percentile of this distribution (i.e., cluster-level threshold $p < .05$, Bonferroni-corrected for three phases of the experiment).

@@

The EEG data are available upon reasonable request from the corresponding author because we had not asked participants for their consent to make the data publicly available.
The experimental stimuli (object images and keywords) the code for data analysis are available at [https://doi.org/10.17605/osf.io/uksbc](https://doi.org/10.17605/osf.io/uksbc).
